Conversations in time: interactive visualization to explore structured temporal data

### 2020-07-28 tsibbletalk.Rmd
draft shiny module

---
title: "Conversations in time: interactive visualization to explore structured temporal data"
author:
- name: Earo Wang
  affiliation: The Univeristy of Auckland
  address:
  - Department of Statistics
  - The Univeristy of Auckland
  - New Zealand
  email: earo.wang@auckland.ac.nz
- name: Dianne Cook
  affiliation: Monash Univeristy
  address:
  - Department of Econometrics and Business Statistics
  - Monash University
  - Australia
  email: dicook@monash.edu
bibliography: references.bib
abstract: >
  An abstract of less than 150 words.
preamble: |
  % Any extra LaTeX you need in the preamble
output: rticles::rjournal_article
---

@@

The other critical aspect of a tsibble is "index", that provides foundational temporal context. A common tool in time series analytical toolkit is seasonal plots that lay time series not on the whole time scale, but on an origin-less relative time unit, such as `gg_season()` in the {feasts} package. It helps to examine and emphasise periodic/aperiodic patterns, comparing to time series plots that primarily focus on trends. Standard seasonal plots break the overall time into two components: seasonal periods on the x-axis, and grouped by their corresponding lower-resolution time. For example, monthly data can be decomposed into months separated by years, and hourly data into hours grouped by days.

* Transform and send newly-transformed data via server, without updating anything else (e.g. layout and graphical elements), to avoid completely redrawing.
* A shiny module: slider and plot

>>>>

---
title: "Conversations in time: interactive visualization to explore structured temporal data"
author:
- name: Earo Wang
  affiliation: The Univeristy of Auckland
  address:
  - Department of Statistics
  - The Univeristy of Auckland
  - New Zealand
  email: earo.wang@auckland.ac.nz
- name: Dianne Cook
  affiliation: Monash Univerisity
  address:
  - Department of Econometrics and Business Statistics
  - Monash University
  - Australia
  email: dicook@monash.edu
bibliography: references.bib
abstract: >
  An abstract of less than 150 words.
preamble: |
  % Any extra LaTeX you need in the preamble
output: rticles::rjournal_article
---

@@

The other critical aspect of a tsibble is "index", that provides foundational temporal context. A common tool in time series analytical toolkit is seasonal plots that lay time series not on the whole time scale, but on an origin-less relative time unit, for example `gg_season()` in the {feasts} package. It helps to examine and emphasise periodic/aperiodic patterns, comparing to time series plots that primarily focus on trends. Standard seasonal plots break the overall time into two components: seasonal periods on the x-axis, and grouped by their corresponding lower-resolution time. For example, monthly data can be decomposed into months separated by years, and hourly data into hours grouped by days. Data collected at lower-level resolutions often exhibits more than one seasonal patterns. To discover typical seasonal or non-typical profiles, it is helpful to quickly browse through many possible periods. Interactivity ought to be enabled.

The {tsibbletalk} package provides a pair of UI and server functions, as a shiny module, to help with finding interesting time slices in a shiny application. The pair, `tsibbleDiceUI()` and `tsibbleDiceServer()`, presents a clean interface and forms a resusable piece. Like all shiny modules, users should supply a unique session id. The UI function `tsibbleDiceUI()` shows a slider that controls the number of periods, and a plot specified by users. The server function `tsibbleDiceServer()` is the workhorse, transforming data and updating the plot. It expects a `ggplot` (converted to `plotly` via `ggplotly()`) or `plotly` object. This plot can be line charts, or other graphical elements (such as boxplots). But it assumes that tsibble's time index is plotted on the x-axis. The other mandatory argument is to specify the number of seasonal periods that requires shifting.

* Transform and send newly-transformed data via server, without updating anything else (e.g. layout and graphical elements), to avoid completely redrawing. Performance.
* Double dispatch

### 2020-07-28 tsibbletalk.tex
draft shiny module

The other critical aspect of a tsibble is ``index'', that provides
foundational temporal context. A common tool in time series analytical
toolkit is seasonal plots that lay time series not on the whole time
scale, but on an origin-less relative time unit, such as
\texttt{gg\_season()} in the \{feasts\} package. It helps to examine and
emphasise periodic/aperiodic patterns, comparing to time series plots
that primarily focus on trends. Standard seasonal plots break the
overall time into two components: seasonal periods on the x-axis, and
grouped by their corresponding lower-resolution time. For example,
monthly data can be decomposed into months separated by years, and
hourly data into hours grouped by days.

\begin{itemize}
\tightlist
\item
  Transform and send newly-transformed data via server, without updating
  anything else (e.g.~layout and graphical elements), to avoid
  completely redrawing.
\item
  A shiny module: slider and plot
\end{itemize}

@@

\address{%
Dianne Cook\\
Monash Univeristy\\
Department of Econometrics and Business Statistics\\ Monash
University\\ Australia\\
}
\href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}

>>>>

The other critical aspect of a tsibble is ``index'', that provides
foundational temporal context. A common tool in time series analytical
toolkit is seasonal plots that lay time series not on the whole time
scale, but on an origin-less relative time unit, for example
\texttt{gg\_season()} in the \{feasts\} package. It helps to examine and
emphasise periodic/aperiodic patterns, comparing to time series plots
that primarily focus on trends. Standard seasonal plots break the
overall time into two components: seasonal periods on the x-axis, and
grouped by their corresponding lower-resolution time. For example,
monthly data can be decomposed into months separated by years, and
hourly data into hours grouped by days. Data collected at lower-level
resolutions often exhibits more than one seasonal patterns. To discover
typical seasonal or non-typical profiles, it is helpful to quickly
browse through many possible periods. Interactivity ought to be enabled.

The \{tsibbletalk\} package provides a pair of UI and server functions,
as a shiny module, to help with finding interesting time slices in a
shiny application. The pair, \texttt{tsibbleDiceUI()} and
\texttt{tsibbleDiceServer()}, presents a clean interface and forms a
resusable piece. Like all shiny modules, users should supply a unique
session id. The UI function \texttt{tsibbleDiceUI()} shows a slider that
controls the number of periods, and a plot specified by users. The
server function \texttt{tsibbleDiceServer()} is the workhorse,
transforming data and updating the plot. It expects a \texttt{ggplot}
(converted to \texttt{plotly} via \texttt{ggplotly()}) or
\texttt{plotly} object. This plot can be line charts, or other graphical
elements (such as boxplots). But it assumes that tsibble's time index is
plotted on the x-axis. The other mandatory argument is to specify the
number of seasonal periods that requires shifting.

\begin{itemize}
\tightlist
\item
  Transform and send newly-transformed data via server, without updating
  anything else (e.g.~layout and graphical elements), to avoid
  completely redrawing. Performance.
\item
  Double dispatch
\end{itemize}

@@

\address{%
Dianne Cook\\
Monash Univerisity\\
Department of Econometrics and Business Statistics\\ Monash
University\\ Australia\\
}
\href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}
